---
title: "Exam 1"
code-fold: false
---

```{r}
library(tidytuesdayR)
library(tidyverse)
library(rnaturalearth)

fc <- tt_load('2020-02-18')$food_consumption
```

## Grand Research Question

What does the consumption of each food category in each country look like?

## Understanding the Data

```{r}
# Let's find out how many observations/variables we're working with
dim(fc)

# Looking at information on variables
names(fc)

str(fc)

# What are the categories of our categorical variables
unique(fc$country)

unique(fc$food_category)

# Peek at the dataset
head(fc)
```

Our observations appear to be groups of food category and country. That means that for every country there is 11 observations for each food group.

## Variable Exploration

```{r}
# Removing variable that is not pertinent to our investigation
fc <- select(fc, -co2_emmission)
```

**Individual variable exploration (not including country or food_cat for obvious reasons)**

```{r}
fc |>
  ggplot(aes(x = consumption)) +
  geom_histogram() +
  labs(x = "Food Consumption (kg/person/year)")
```

From the histogram, we can see that it is most common for food to be consumed at a rate between 0 and 100 kilograms per person per year, however, consumption rates can reach as high as 350 kilograms per person per year.

**Exploring relationships between variables**

```{r}
# Consumption by food category
fc |>
  ggplot(aes(x = food_category, y = consumption)) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

# Consumption by country
fc |>
  ggplot(aes(x = consumption)) +
  geom_histogram() +
  facet_wrap(~country) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```

From the visualizations, we can see that the most consumed food category is milk and cheese, the second highest is wheat and wheat derivatives.
The lowest seems to be soybeans and then lamb and goat.
The second visualization is too difficult to interpret.

## Answering Grand RQ

```{r}
#| fig-height: 15
#| fig-width: 40
#| fig-alt: "A group of histograms, facet-wrapped by food category, that depict the level of consumption in term of kilograms per person per year in 130 countries. Difficult to distinguish any noticeable patterns."

fc |>
  ggplot(aes(x = country, y = consumption)) +
  geom_boxplot() +
  facet_wrap(~food_category) +
  theme(axis.text.x = element_text(
    angle = 45, 
    vjust = 1, 
    hjust = 1,
    size = 5)) +
  labs(x = "Country", 
       y = "Food Consumption (kg/person/year)",
       title = "Food Consumption by category of food for each country",
       caption = "Visualization by Marvin C. (2025), data sourced from TidyTuesday's February 2020 dataset on food consumption")
```

It is difficult to format this, but an interesting thing to note is that Finland has the highest consumption of milk and cheese, with nearly 400 kilograms per person per year. Countries of lower economic status tend to have lower consumption overall, the most consumed food cate
