---
title: "Exam 2"
format: html
---

```{r}
library(tidytuesdayR)
library(tidyverse)
library(rnaturalearth)
library(sf)
library(viridis)
```

```{r}
fc <- tt_load('2020-02-18')$food_consumption
```

## Inspecting the Data

```{r}
# What variables should we expect to be working with?
str(fc)
```

```{r}
# What does our dataset look like?
head(fc, 22)
```

The observations in the data seem like they're countries and food groups, so an observation for each possible combination of food category and country.

**Inspecting individual variables**

```{r}
# What are our categories?
distinct(fc, country)

distinct(fc, food_category)
```

There are 130 unique countries and 11 unique food categories. One issue that may arise when creating visualizations is that the food category names are too long and specific, so we should shorten them without loss of information.

**Fixing the food_category variable**

```{r}
# Shortening longer variable names
fcc <- fc |>
  mutate(food_category = str_replace(food_category, " & Goat", ""),
         food_category = str_replace(food_category, "Milk - inc. cheese", "Dairy"),
         food_category = str_replace(food_category, "Wheat and Wheat Products", "Wheat"),
         food_category = str_replace(food_category, "Nuts inc. Peanut Butter", "Nuts"))
```

```{r}
# Assuring relabeled values appear as expected
fcc |> distinct(food_category)
```

## Countries with the Highest Consumption

**Research Question 1:** Which 5 countries consume the most food?

```{r}
# Wrangling cleaned dataset to answer question
fcc |>
  group_by(country) |>
  summarize(total_consumption = sum(consumption)) |>
  arrange(desc(total_consumption)) |>
  head(5)
```
 
The top 5 countries in terms of food consumption are, in order from greatest to least, Finland, Lithuania, Sweden, Netherlands, and Albania.

```{r}
#| fig-alt: "A column chart depicting the top 5 countries in the world in terms of their total food consumption. The top 5 countries, in order from least to greatest, are Albania, Netherlands, Sweden, Lithuania, and Finland. The values range from 532.73 to 639.79 kilograms per person per year"

# Visualizing the wrangled dataset
fcc |>
  group_by(country) |>
  summarize(total_consumption = sum(consumption)) |>
  arrange(desc(total_consumption)) |>
  head(5) |>
  ggplot(aes(x = country, y = total_consumption)) +
  geom_col() +
  scale_y_continuous(breaks = seq(500, 650, 50)) +
  coord_cartesian(ylim = c(500, 650)) +
  labs(x = "Country", 
       y = "Total Consumption (kg/person/year)", 
       title = "Top 5 Countries by Total Food Consumption",
       caption = "Visualization by Marvin C. (2025), data sourced from TidyTuesday's February 2020 dataset on global food consumption")
```

## Most Consuming Countries of Each Food Category

**Research Question 2:** What are the top 5 countries in terms of consumption of each food group?

```{r}
# Wrangling dataset to answer our question
fcc |>
  group_by(food_category, country) |>
  summarize(total_consumption = sum(consumption)) |>
  slice_max(n = 5, order_by = total_consumption)
```

```{r}
#| fig-alt: "A group of column charts where each column chart represents a different food group and depicts the top 5 countries that consume that food group. No particular trends to note, other than the fact that dairy, wheat, and rice are consumed at much greater than other food groups."

fcc |>
  group_by(food_category, country) |>
  summarize(total_consumption = sum(consumption)) |>
  slice_max(n = 5, order_by = total_consumption) |>
  ggplot(aes(x = country, 
             y = total_consumption)) +
  geom_col(width = 0.3) +
  facet_wrap(~food_category,
             ncol = 4,
             scales = "free_x") +
  theme(axis.text.x = element_text(angle = 25, vjust = 1.1, hjust=0.8)) +
  labs(x = "Country", 
       y = "Total Consumption (kg/person/year)", 
       title = "Top 5 Countries Consuming each Food Group",
       caption = "Visualization by Marvin C. (2025), data sourced from TidyTuesday's February 2020 dataset on global food consumption")
```

The graphs and table both answer our question.

## Food Consumption Map

**Research Question 3:** What does the consumption of each food look like?

```{r}
# Loading necessary information to answer question
world <- ne_countries(returnclass = "sf")
```

**Viz Attempt 1**

```{r}
world |>
  select(name, geometry) |>
  left_join(fcc |> select(-co2_emmission),
            join_by(name == country)) |>
  ggplot() +
  geom_sf(aes(fill = consumption)) +
  facet_wrap(~food_category) +
  theme(legend.position = "bottom")
```

Good first attempt. Some issues: 

\- NA should not be represented

\- We are missing actual data on some countries, like the US and a good chunk of Africa

Let's find out what countries don't have values associated with them for our next attempt 

```{r}
world |>
  select(name) |>
  anti_join(fcc, join_by(name == country)) |>
  st_drop_geometry() |>
  arrange(name)
```

**Viz Attempt 2**

```{r}
world |>
  select(name, geometry) |>
  mutate(name = ifelse(name == "United States of America", "USA", name)) |>
  mutate(name = ifelse(name == "Bosnia and Herz.", "Bosnia and Herzegovina", name)) |>
  mutate(name = ifelse(name == "Czechia", "Czech Republic", name)) |>
  mutate(name = ifelse(name == "Taiwan", "Taiwan. ROC", name)) |>
  left_join(fcc |> select(-co2_emmission),
            join_by(name == country)) |>
  pivot_wider(names_from = food_category,
              values_from = consumption) |>
  select(-"NA") |>
  pivot_longer(cols = c(-name, -geometry),
               names_to = "food_category",
               values_to = "consumption") |>
  ggplot() +
  geom_sf(aes(fill = consumption)) +
  facet_wrap(~food_category) +
  theme(legend.position = "bottom")
```

This is a better visualization, but there is still an issue: the color scale makes it so that instead of looking at the consumption of each food category *individually*, we see them in relation to each other. We can solve this by instead taking the averages of the consumption of each food category, and standardizing the scale. That way we can see if each country consumes an average, more than average, or less than average amount of that food group.

First, let's look at some numerical summaries to get an idea of this: 

```{r}
fcc |>
  group_by(food_category) |>
  summarize(min = min(consumption), 
            max = max(consumption), 
            range = max - min, 
            mean = mean(consumption, na.rm = TRUE),
            sd = sd(consumption, na.rm = TRUE))
```

**Final Viz Attempt**

```{r}
#| fig-alt: "A set of chloropeth maps facet-wrapped by 11 food categories. The fill color represents the average consumption of that food group throughout the world, and we can see that dairy is by far the most consumed product in the world, especially in Europe and America. Wheat is trailing 2nd in the world."
world |>
  select(name, geometry) |>
  mutate(name = ifelse(name == "United States of America", "USA", name)) |>
  mutate(name = ifelse(name == "Bosnia and Herz.", "Bosnia and Herzegovina", name)) |>
  mutate(name = ifelse(name == "Czechia", "Czech Republic", name)) |>
  mutate(name = ifelse(name == "Taiwan", "Taiwan. ROC", name)) |>
  left_join(
    fcc |>
      select(-co2_emmission) |>
      group_by(food_category) |>
      mutate(consumption = consumption - (mean(consumption) / sd(consumption))),
    join_by(name == country)) |>
  pivot_wider(names_from = food_category, values_from = consumption) |>
  select(-"NA") |>
  pivot_longer(cols = c(-name, -geometry),
               names_to = "food_category",
               values_to = "consumption") |>
  ggplot() +
  geom_sf(aes(fill = consumption)) +
  facet_wrap(~food_category) +
  theme(legend.position = "bottom") +
  scale_fill_viridis() +
  labs(title = "Consumption of Different Food Categories Throughout the World",
       caption = "Visualization created by Marvin C. (2025), data sourced from TidyTuesday's February 2020 dataset on food consumption")
```

