---
title: "Solo Project"
format: html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r}
library(tidyverse)
library(rnaturalearth)

states <- ne_states(country = "united states of america", returnclass = 'sf')
```

```{r}
congress <- read_csv("https://raw.githubusercontent.com/fivethirtyeight/data/refs/heads/master/congress-demographics/data_aging_congress.csv")

congress_bipart <- congress |>
  filter(party_code %in% c(100, 200), # Only republican (200) and democrat (100) congresspeople
         congress == 118, # Viewing data on members of the 118th (most recent before current) congress
         !(state_abbrev %in% c("HI", "AK"))) |>  # Excluding Hawaii and Alaska for a cleaner map
  mutate(party = if_else(party_code == 100, "dem", "rep")) |> # Adding party labels
  select(-party_code, -start_date, -bioguide_id, -cmltv_cong, -cmltv_chamber, -congress, -age_days) # Pruning unnecessary data
```

```{r}
senate_bipart <- congress_bipart |> 
  filter(chamber == "Senate") |> # only senators 
  select(-chamber)

house_bipart <- congress_bipart |> 
  filter(chamber == "House") |> # only members of the house 
  select(-chamber)
```

```{r}
# Summaries of ages of congress members by state
congress_bipart |>
  group_by(state_abbrev) |>
  summarize(total = n(),
            mean_age = mean(age_years),
            max_age = max(age_years),
            min_age = min(age_years),
            median_age = median(age_years))

# Summaries of ages of congress members by party
congress_bipart |>
  group_by(party) |>
  summarize(total = n(),
            mean_age = mean(age_years),
            max_age = max(age_years),
            min_age = min(age_years),
            median_age = median(age_years))

# Age summaries of the senate by party
senate_bipart |>
  group_by(party) |>
  summarize(total = n(),
            mean_age = mean(age_years),
            max_age = max(age_years),
            min_age = min(age_years),
            median_age = median(age_years))

# Age summaries of the house by party
house_bipart |>
  group_by(party) |>
  summarize(total = n(),
            mean_age = mean(age_years),
            max_age = max(age_years),
            min_age = min(age_years),
            median_age = median(age_years))
```

```{r}
dem_ages_by_state <- congress_bipart |> # Average ages of Democrat congress members by state
  filter(party == "dem") |>
  group_by(state_abbrev) |>
  summarize(dem_avg_age = mean(age_years))

rep_ages_by_state <- congress_bipart |> # Average ages of Republican congress members by state
  filter(party == "rep") |>
  group_by(state_abbrev) |>
  summarize(rep_avg_age = mean(age_years))

all_ages_by_state <- congress_bipart |> # Average ages of all congress members by state
  group_by(state_abbrev) |>
  summarize(avg_age = mean(age_years))
```


```{r}
#|fig-alt: "A chloropleth map of the contiguous United States depicting the average ages of all congress members of the 118th Congress, active from 2023 to 2025. The map shows that the oldest congressmembers on average work in the Northwest portion of the United States, and in the upper parts of New England. The states with lowest average ages are New Mexico, New York, Colorado, and Kansas."

ggplot() +
  geom_sf(data = states |> 
            filter(!(postal %in% c("HI", "AK"))) |>
            left_join(all_ages_by_state, join_by("postal" == "state_abbrev")),  # Contiguous US
          aes(fill = avg_age),
          color = 'white') +
  scale_fill_gradient(low = 'lavenderblush2', high = 'orchid4') +
  theme_void() +
  labs(title = "Average Age of Congress Members in the 118th Congress by State", fill = "Average Age (yrs)", caption = "Visualization by Marvin C. 2025, data sourced from the fivethirtyeight Github repository.")
```

```{r}
#|fig-alt: "A chloropleth map of the contiguous United States depicting the average age of Democrats in the 118th Congress, which was active from 2023 to 2025. The states of North and South Dakota, Iowa, Idaho, Utah, Wyoming, Kansas, Oklahoma, Arkansas have no data associated with them. The highest average age by fair is in South Carolina, with the South generally having higher averages ages overall. The lowest averages seem to be in Kansas and Kentucky."

ggplot() +
  geom_sf(data = states |> 
            left_join(dem_ages_by_state, join_by("postal" == "state_abbrev")) |>
            filter(!(postal %in% c("HI", "AK"))), # Contiguous US
          aes(fill = dem_avg_age),
          color = 'white') +
  scale_fill_gradient(low = 'lightsteelblue1', high = 'royalblue4') +
  theme_void() +
  labs(title = "Average Age of Congress Democrats in the 118th Congress by State", fill = "Average Age (yrs)", caption = "Visualization by Marvin C. 2025, data sourced from the fivethirtyeight Github repository.")
```

```{r}
#|fig-alt: "A chloropleth map of the contiguous United States depicting the average age of Republicans in the 118th Congress, which was active from 2023 to 2025. The states of Vermont, Massachusetts, Rhode Island, New Mexico, Connecticut, New Hampshire, and Delaware have no data associated with them. The highest average ages are in Maine and Idaho, with the highest average age being found in the pacific northwest and the lowest being in New York."

ggplot() +
  geom_sf(data = states |> 
            left_join(rep_ages_by_state, join_by("postal" == "state_abbrev")) |>
            filter(!(postal %in% c("HI", "AK"))), # Contiguous US
          aes(fill = rep_avg_age),
          color = 'white') +
  scale_fill_gradient(low = 'lightpink', high = 'firebrick4') +
  theme_void() +
  labs(title = "Average Age of Congress Republicans in the 118th Congress by State", fill = "Average Age (yrs)", caption = "Visualization by Marvin C. 2025, data sourced from the fivethirtyeight Github repository.")
```

# Summary

The maps above represent the average ages of the congress members of the 118th congress with individual representations by party and as a whole. The maps tell us that the states which contain only Republican congress members have a higher average age. Similarly, overall the overall average age of congress members is in the range of 50-60 years. Averages above this range are less likely by state. Also there are fewer states that have Democratic congress members than there are states that have Republican congress members. 
